\textsc{PAPERCLIP}: Associating Astronomical Observations and Natural Language with Multi-Modal Models

### 2024-03-03 paper/main.tex
merge

\usepackage{url}
\usepackage{graphicx}
\usepackage{array}
\usepackage{booktabs}
\usepackage{lipsum}
\usepackage{xspace}
\usepackage{pdflscape} 
\usepackage{multirow} 
\usepackage{multicol}
\usepackage{listings}

@@

\begin{figure*}[!h]
\includegraphics[width=0.95\textwidth]{plots/tti_base.pdf}
\caption{Image retrieval using the base CLIP model on four curated queries. \SM{Horizontal lines} \SM{Query not rotated} \SM{Same for next Figure.} \SM{Put text ``Base Model'' in consistent colour to legend.} \SM{Say top 4} \SM{Say that for science we care about the distributions of things.}}
\label{fig:tti_base}
\end{figure*}

@@

\begin{figure*}[!h]
\includegraphics[width=0.95\textwidth]{plots/itt.pdf}
\caption{Text associations from a curated list most closely matching a given image query, for both the fine-tuned and base models. The `ground truth' LLM-summarized abstract is shown in the right column. \SM{Say ground truth summarized abstract in the column}\SM{Swap order of cols 2 and 3.}}
\label{fig:itt}
\end{figure*}

@@

\subsection{Generation of single-concept summaries}
\label{app:singleconcept}

The following prompt is used to generate a list of diverse single-concept summaries informed by a list of summarized abstracts:

@outlines.prompt
def prompt_fn(objects):
    """<s>[INST] Please produce a list of around concepts characterizing prominent objects, phenomena, and science use cases of images observed by the Hubble Space Telescope.

Here are some examples of objects:

{{objects}}

Follow these instructions exactly in your answer:
- Do not output empty strings as elements.
- Make sure that the list covers a diverse range of astronomical concepts, with items as different from each other as possible. 
- Do not give specific names of objects, to make sure you span the widest possible range of concepts (e.g., "dwarf galaxy" is allowed, but NOT "Fornax", "Terzan 5", or  "NGC6440").
- Do not return terms undescriptive of observations, e.g. "sloshing", "adiabatic", "interactions". Returning concrete physics objects, concepts, or phenomena.
- Only output scientifically meaningful terms. E.g., NO "Cosmic Dance".
- Do not duplicate entries. Do not reference any telescopes, observatories, or surveys.
- Do not include units like "angular diameter distance", "parsec", or any other concepts that will not correlate with images of observations.
- Use the above example list of objects only as inspiration to infer broad classes of objects.
- Make sure each concept is succint, never more than 5 words.
- Answer in JSON format.
- The JSON should have the following keys {"galaxies", "stellar_physics", "exoplanets_planet_formation", "stellar_populations", "supermassive_black_holes", "solar_system", "integalactic_medium", "large_scale_structure"} reflecting rough observation categories.
- Each category will have a list of objects and/or astronomical concepts.
- Output up to 20 items and no more in each category
[/INST]
"""
\end{lstlisting}

The following schema guides generation, intended to produce 100 concepts reflecting the science categories of successful proposals\footnote{\url{https://www.stsci.edu/contents/newsletters/2023-volume-40-issue-02/hubble-cycle-31-proposal-selection}}.
%

class ScienceCategoriesHST(BaseModel):
    galaxies: conlist(str, min_length=15, max_length=15)
    stellar_physics: conlist(str, min_length=15, max_length=15)
    exoplanets_planet_formation: conlist(str, min_length=15, max_length=15)
    stellar_populations: conlist(str, min_length=10, max_length=10)
    supermassive_black_holes: conlist(str, min_length=15, max_length=15)
    solar_system: conlist(str, min_length=10, max_length=10)
    integalactic_medium: conlist(str, min_length=10, max_length=10)
    large_scale_structure: conlist(str, min_length=10, max_length=10)
  \end{lstlisting}

The output is then constrained to be one of the fixed number of concepts.

\subsection{Assignment of abstracts to single-concept categories}
\label{app:singleconceptassignments}

Finally, the following prompt is used to assign concepts inferred in \ref{app:singleconcept} to each abstract:

@outlines.prompt
def prompt_fn(abs, cats):
    """<s>[INST] The following is a successful proposal abstract for the Hubble Space Telescope: "{{abs}}"

The following is a list of categories (astronomical concepts) that this abstract could correspond to.

{{cats}}

Please answer which of these listed concepts best describes this proposal, based on the objects and phenomena mentioned in the abstract.
The concept should meaningfully be present in the abstract and the eventual observation.

- For example, "The locations of supernovae {SNe} in the local stellar and gaseous environment in galaxies, as measured in high spatial resolution WFPC2 and ACS images, contain important clues to their progenitor stars." should return "supernova".
- If the abstract centers calibration and/or instrumentation efforts, return calibration or instrumention".

If no concept make sense, return "None". [/INST]
"""
\end{lstlisting}

@@

The following curated categories are used in the text retrieval experiment in Sec.~\ref{sec:results}.
%
These are intended to be broader than the single-concept summaries inferred in Sec.~\ref{app:singleconcept} and are derived by prompting \textsc{Claude}, without any external input e.g.
%
proposal abstracts, to produce a list of categories corresponding to HST observations.

>>>>

\usepackage{url}
\usepackage{graphicx}
\usepackage{array}
\usepackage{booktabs}
\usepackage{lipsum}
\usepackage{xspace}
\usepackage{pdflscape} 
\usepackage{multirow} 
\usepackage{multicol}
\usepackage{listings}
\usepackage{verbatim}

@@

\begin{table}[h!]
  \centering
  \begin{tabular}{m{3cm} p{3cm} p{3cm} p{3cm} p{3cm}}
      \toprule
      \centering \bfseries Query & \multicolumn{4}{c}{\bfseries{Top-4 most similar images}} \tabularnewline
      \midrule
       \texttt{\input{\thedatafolder/query_tti_base_0.txt}} \vspace{20mm} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_0_0.pdf} \\ \input{\thedatafolder/propid_tti_base_0_0.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_0_1.pdf} \\ \input{\thedatafolder/propid_tti_base_0_1.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_0_2.pdf} \\ \input{\thedatafolder/propid_tti_base_0_2.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_0_3.pdf} \\ \input{\thedatafolder/propid_tti_base_0_3.txt}  \tabularnewline
      \midrule
      \texttt{\input{\thedatafolder/query_tti_base_1.txt}} \vspace{20mm} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_1_0.pdf} \\ \input{\thedatafolder/propid_tti_base_1_0.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_1_1.pdf} \\ \input{\thedatafolder/propid_tti_base_1_1.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_1_2.pdf} \\ \input{\thedatafolder/propid_tti_base_1_2.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_1_3.pdf} \\ \input{\thedatafolder/propid_tti_base_1_3.txt}  \tabularnewline
      \midrule
      \texttt{\input{\thedatafolder/query_tti_base_2.txt}} \vspace{20mm} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_2_0.pdf} \\ \input{\thedatafolder/propid_tti_base_2_0.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_2_1.pdf} \\ \input{\thedatafolder/propid_tti_base_2_1.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_2_2.pdf} \\ \input{\thedatafolder/propid_tti_base_2_2.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_2_3.pdf} \\ \input{\thedatafolder/propid_tti_base_2_3.txt}  \tabularnewline
      \midrule
      \texttt{\input{\thedatafolder/query_tti_base_3.txt}} \vspace{20mm} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_3_0.pdf} \\ \input{\thedatafolder/propid_tti_base_3_0.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_3_1.pdf} \\ \input{\thedatafolder/propid_tti_base_3_1.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_3_2.pdf} \\ \input{\thedatafolder/propid_tti_base_3_2.txt} & \centering \includegraphics[width=0.2\textwidth]{\thedatafolder/img_tti_base_3_3.pdf} \\ \input{\thedatafolder/propid_tti_base_3_3.txt}  \tabularnewline
      \bottomrule
  \end{tabular}
  \caption{Image retrieval using the base CLIP model on four curated queries. \SM{Horizontal lines} \SM{Query not rotated} \SM{Same for next Figure.} \SM{Put text ``Base Model'' in consistent colour to legend.} \SM{Say top 4} \SM{Say that for science we care about the distributions of things.}}
  \label{tab:tti_base}
\end{table}

% \begin{figure*}[!h]
% \includegraphics[width=0.95\textwidth]{plots/tti_base.pdf}
% \caption{Image retrieval using the base CLIP model on four curated queries. \SM{Horizontal lines} \SM{Query not rotated} \SM{Same for next Figure.} \SM{Put text ``Base Model'' in consistent colour to legend.} \SM{Say top 4} \SM{Say that for science we care about the distributions of things.}}
% \label{fig:tti_base}
% \end{figure*}

@@

% \begin{figure*}[!h]
% \includegraphics[width=0.95\textwidth]{plots/itt.pdf}
% \caption{Text associations from a curated list most closely matching a given image query, for both the fine-tuned and base models. The `ground truth' LLM-summarized abstract is shown in the right column. \SM{Say ground truth summarized abstract in the column}\SM{Swap order of cols 2 and 3.}}
% \label{fig:itt}
% \end{figure*}

% Set line spacing to 0.5

\begin{table}[h!]
  \centering
  \renewcommand{\arraystretch}{0.1}
  \begin{tabular}{m{3cm} m{3cm} m{3cm} m{6cm}}
      \toprule
      \centering \bfseries \hubble Image & \centering \bfseries Obs. cycle \\ (Year) & \centering \bfseries Prop. ID & \centering \bfseries Proposal abstract (clipped) \tabularnewline
      \midrule
      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_0.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_0.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_0.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_0.txt}} \tabularnewline
      \midrule
      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_1.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_1.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_1.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_1.txt}} \tabularnewline
      \midrule
      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_2.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_2.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_2.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_2.txt}} \tabularnewline
      \midrule
      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_3.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_3.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_3.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_3.txt}} \tabularnewline
      \bottomrule
  \end{tabular}
  \caption{Text associations from a curated list most closely matching a given image query, for both the fine-tuned and base models. The `ground truth' LLM-summarized abstract is shown in the right column. \SM{Say ground truth summarized abstract in the column}\SM{Swap order of cols 2 and 3.}}
  \label{tab:itt}
\end{table}

@@

% \subsection{Generation of single-concept summaries}
% \label{app:singleconcept}

% The following prompt is used to generate a list of diverse single-concept summaries informed by a list of summarized abstracts:

% \begin{lstlisting}[language=Python]
% import outlines

% @outlines.prompt
% def prompt_fn(objects):
%     """<s>[INST] Please produce a list of around concepts characterizing prominent objects, phenomena, and science use cases of images observed by the Hubble Space Telescope.

% Here are some examples of objects:

% {{objects}}

% Follow these instructions exactly in your answer:
% - Do not output empty strings as elements.
% - Make sure that the list covers a diverse range of astronomical concepts, with items as different from each other as possible. 
% - Do not give specific names of objects, to make sure you span the widest possible range of concepts (e.g., "dwarf galaxy" is allowed, but NOT "Fornax", "Terzan 5", or  "NGC6440").
% - Do not return terms undescriptive of observations, e.g. "sloshing", "adiabatic", "interactions". Returning concrete physics objects, concepts, or phenomena.
% - Only output scientifically meaningful terms. E.g., NO "Cosmic Dance".
% - Do not duplicate entries. Do not reference any telescopes, observatories, or surveys.
% - Do not include units like "angular diameter distance", "parsec", or any other concepts that will not correlate with images of observations.
% - Use the above example list of objects only as inspiration to infer broad classes of objects.
% - Make sure each concept is succint, never more than 5 words.
% - Answer in JSON format.
% - The JSON should have the following keys {"galaxies", "stellar_physics", "exoplanets_planet_formation", "stellar_populations", "supermassive_black_holes", "solar_system", "integalactic_medium", "large_scale_structure"} reflecting rough observation categories.
% - Each category will have a list of objects and/or astronomical concepts.
% - Output up to 20 items and no more in each category
% [/INST]
% """
% \end{lstlisting}

% The following schema guides generation, intended to produce 100 concepts reflecting the science categories of successful proposals\footnote{\url{https://www.stsci.edu/contents/newsletters/2023-volume-40-issue-02/hubble-cycle-31-proposal-selection}}.
% %

% \begin{lstlisting}[language=Python]
% from pydantic import BaseModel, conlist

% class ScienceCategoriesHST(BaseModel):
%     galaxies: conlist(str, min_length=15, max_length=15)
%     stellar_physics: conlist(str, min_length=15, max_length=15)
%     exoplanets_planet_formation: conlist(str, min_length=15, max_length=15)
%     stellar_populations: conlist(str, min_length=10, max_length=10)
%     supermassive_black_holes: conlist(str, min_length=15, max_length=15)
%     solar_system: conlist(str, min_length=10, max_length=10)
%     integalactic_medium: conlist(str, min_length=10, max_length=10)
%     large_scale_structure: conlist(str, min_length=10, max_length=10)
%   \end{lstlisting}

% The output is then constrained to be one of the fixed number of concepts.

% \subsection{Assignment of abstracts to single-concept categories}
% \label{app:singleconceptassignments}

% Finally, the following prompt is used to assign concepts inferred in \ref{app:singleconcept} to each abstract:

% \begin{lstlisting}[language=Python]
% import outlines

% @outlines.prompt
% def prompt_fn(abs, cats):
%     """<s>[INST] The following is a successful proposal abstract for the Hubble Space Telescope: "{{abs}}"

% The following is a list of categories (astronomical concepts) that this abstract could correspond to.

% {{cats}}

% Please answer which of these listed concepts best describes this proposal, based on the objects and phenomena mentioned in the abstract.
% The concept should meaningfully be present in the abstract and the eventual observation.

% - For example, "The locations of supernovae {SNe} in the local stellar and gaseous environment in galaxies, as measured in high spatial resolution WFPC2 and ACS images, contain important clues to their progenitor stars." should return "supernova".
% - If the abstract centers calibration and/or instrumentation efforts, return calibration or instrumention".

% If no concept make sense, return "None". [/INST]
% """
% \end{lstlisting}

@@

The following curated categories are used in the text retrieval experiment in Sec.~\ref{sec:results}.
%
These are derived by prompting \textsc{Claude}, without any external input e.g.
%
proposal abstracts, to produce a list of categories corresponding to HST observations.
